Graphs

### 2023-11-11 graphs/paper.tex
graphs

\begin{rmk}
  \label{rmk:edges-are-binary-relations}
  Recall that a binary relation on a set $V$ can be encoded as any subset of $V \times V$.
  Thus, a graph is a set $V$ equipped with a binary relation $E$.
\end{rmk}

@@

\begin{rmk}
  \label{rmk:example1-not-undirected}
  \cref{eg:graphs-example1} is not an undirected graph since $(B,C) \in E$, but $(C,B) \notin E$.
  Similarly, $(C,A) \in E$, but $(A,C) \notin E$.
\end{rmk}

\begin{rmk}
  \label{rmk:drop-arrows-in-undirected-graphs}
  In an undirected graph, we drop the arrow tips as they convey no additional information.
\end{rmk}

@@

\begin{thm}
  \label{thm:reachability-is-an-equivalence}
  Let $G = (V, E)$ be an undirected graph.
  The s-t reachability relation defined in \cref{defn:reachability} is an equivalence relation on the set of vertices $V$.
\end{thm}
\begin{proof}
  (Reflexivity): Let $v \in V$.
  The empty walk is trivially a path.
  Thus, $v$ is s-t reachable from itself.

  (Symmetry): Let $u,v \in V$ be any two vertices.
  Assume that $v$ is s-t reachable from $u$, i.e., there is a path $p$ from $u$ to $v$.
  We need to prove that $u$ is s-t reachable from $v$, i.e., there is a path from $v$ to $u$.
  To this end, we do induction on the length of $p$.
  In the base case, $p$ has length 0, so it is the empty path.
  Thus, the result follows from reflexivity.
  In the induction step, $p$ is a path of length $k+1$ and it is of the form $p',(u',v),v$, where $u'$ is some vertex and $p'$ is a path from $u$ to $u'$ of length $k$.
  By the induction hypothesis, there is a path $\widetilde{p}$ from $u'$ to $u$.
  Since $G$ is undirected, $(v,u')$ is also an edge of $G$.
  Thus, the sequence
  \[
    v,(v,u'),\widetilde{p}
  \]
  defines a path from $v$ to $u$.

  (Transitivity): Let $u,v,w \in V$ be any three vertices.
  Assume that there is a path $p$ from $u$ to $v$ and a path $p'$ from $v$ to $w$.
  The concatenation of $p$ and $p'$ defines a path from $u$ to $w$.
  You can prove this rigorously by induction on the length of $p$.
  The detail is left as an exercise.
\end{proof}

@@

\begin{rmk}
  \label{rmk:graph-exploration-algorithms-main-idea-1}
  To explore every vertex of an undirected graph $(V,E)$, we divide $V$ into two piles: frontier and unexplored.
\end{rmk}

\begin{rmk}
  \label{rmk:graph-exploration-algorithms-main-idea-2}
  If the unexplored pile is empty, then that means every vertex has been explored.
  We can remove everything from the frontier pile.
\end{rmk}

\begin{rmk}
  \label{rmk:graph-exploration-algorithms-main-idea-3}
  If the frontier pile is empty, move an arbitrary vertex in the unexplored pile to the frontier pile.
\end{rmk}

\begin{rmk}
  \label{rmk:graph-exploration-algorithms-main-idea-4}
  If the frontier pile is not empty, pick the next vertex $v$ in the frontier pile and move every vertex adjacent to $v$ in the unexplored pile to the frontier pile.
  At this point, $v$ has been fully explored, so we can remove $v$ from the frontier pile.
\end{rmk}

\begin{rmk}
  \label{rmk:graph-exploration-algorithms-next-vertex}
  The ``next vertex'' in \cref{rmk:graph-exploration-algorithms-main-idea-4} is intentionally abstract.
  There are two natural ways to pick the ``next vertex:'' pick the first vertex or the last vertex added to the frontier pile.
  The former choice corresponds to breadth-first search (BFS) and the latter choice corresponds to depth-first choice (DFS).
\end{rmk}

\begin{eg}
  \label{eg:bfs-example}
  Consider the following graph:
  \begin{center}
    \begin{tikzpicture}[node distance= 2cm]
      \begin{scope}[every node/.style={circle,thick,draw}]
        \node (A) {$A$};
        \node (B) [above right of= A] {$B$};
        \node (C) [below right of= A] {$C$};
        \node (D) [right of= B] {$D$};
        \node (E) [right of= C] {$E$};
        \node (F) [below right of= D] {$F$};
      \end{scope}
      \begin{scope}
        [every edge/.style={draw= black, very thick},
        every node/.style={fill= white,circle}]
        \path (A) edge (B);
        \path (A) edge (C);
        \path (B) edge (C);
        \path (B) edge (D);
        \path (B) edge (E);
        \path (C) edge (D);
        \path (C) edge (E);
        \path (D) edge (E);
        \path (D) edge (F);
        \path (E) edge (F);
      \end{scope}
    \end{tikzpicture}
  \end{center}
  We run BFS on this graph.

  (Step 1): The frontier pile is empty and the unexplored pile consists of every vertex in the graph.
  By \cref{rmk:graph-exploration-algorithms-main-idea-3} we move an arbitrary vertex, e.g, $B$, to the frontier pile.
  \begin{center}
    \begin{tikzpicture}[node distance= 2cm]
      \begin{scope}[every node/.style={circle,thick,draw}]
        \node (A) {$A$};
        \node[blue] (B) [above right of= A] {$B$};
        \node (C) [below right of= A] {$C$};
        \node (D) [right of= B] {$D$};
        \node (E) [right of= C] {$E$};
        \node (F) [below right of= D] {$F$};
      \end{scope}
      \begin{scope}
        [every edge/.style={draw= black, very thick},
        every node/.style={fill= white,circle}]
        \path (A) edge (B);
        \path (A) edge (C);
        \path (B) edge (C);
        \path (B) edge (D);
        \path (B) edge (E);
        \path (C) edge (D);
        \path (C) edge (E);
        \path (D) edge (E);
        \path (D) edge (F);
        \path (E) edge (F);
      \end{scope}
    \end{tikzpicture}
  \end{center}

  (Step 2): By \cref{rmk:graph-exploration-algorithms-main-idea-4}, pick the next vertex $B$.
  We have to move $A$, $C$, $D$, and $E$ to the frontier pile and remove $B$ from the frontier pile.
  \begin{center}
    \begin{tikzpicture}[node distance= 2cm]
      \begin{scope}[every node/.style={circle,thick,draw}]
        \node[blue] (A) {$A$};
        \node[gray] (B) [above right of= A] {$B$};
        \node[blue] (C) [below right of= A] {$C$};
        \node[blue] (D) [right of= B] {$D$};
        \node[blue] (E) [right of= C] {$E$};
        \node (F) [below right of= D] {$F$};
      \end{scope}
      \begin{scope}
        [every edge/.style={draw= black, very thick},
        every node/.style={fill= white,circle}]
        \path (A) edge[red] (B);
        \path (A) edge (C);
        \path (B) edge[red] (C);
        \path (B) edge[red] (D);
        \path (B) edge[red] (E);
        \path (C) edge (D);
        \path (C) edge (E);
        \path (D) edge (E);
        \path (D) edge (F);
        \path (E) edge (F);
      \end{scope}
    \end{tikzpicture}
  \end{center}

  (Step 3): By \cref{rmk:graph-exploration-algorithms-main-idea-4}, pick the next vertex $A$.
  We do not have to move anything to the frontier pile because it is not adjacent to any unexplored vertex.
  Finally, we need to remove $A$ from the frontier pile.
  \begin{center}
    \begin{tikzpicture}[node distance= 2cm]
      \begin{scope}[every node/.style={circle,thick,draw}]
        \node[gray] (A) {$A$};
        \node[gray] (B) [above right of= A] {$B$};
        \node[blue] (C) [below right of= A] {$C$};
        \node[blue] (D) [right of= B] {$D$};
        \node[blue] (E) [right of= C] {$E$};
        \node (F) [below right of= D] {$F$};
      \end{scope}
      \begin{scope}
        [every edge/.style={draw= black, very thick},
        every node/.style={fill= white,circle}]
        \path (A) edge[red] (B);
        \path (A) edge[gray] (C);
        \path (B) edge[red] (C);
        \path (B) edge[red] (D);
        \path (B) edge[red] (E);
        \path (C) edge (D);
        \path (C) edge (E);
        \path (D) edge (E);
        \path (D) edge (F);
        \path (E) edge (F);
      \end{scope}
    \end{tikzpicture}
  \end{center}

  (Step 4): By \cref{rmk:graph-exploration-algorithms-main-idea-4}, pick the next vertex $C$.
  Again, we do not have to move anything to the frontier pile because it is not adjacent to any unexplored vertex.
  Finally, we need to remove $C$ from the frontier pile.
  \begin{center}
    \begin{tikzpicture}[node distance= 2cm]
      \begin{scope}[every node/.style={circle,thick,draw}]
        \node[gray] (A) {$A$};
        \node[gray] (B) [above right of= A] {$B$};
        \node[gray] (C) [below right of= A] {$C$};
        \node[blue] (D) [right of= B] {$D$};
        \node[blue] (E) [right of= C] {$E$};
        \node (F) [below right of= D] {$F$};
      \end{scope}
      \begin{scope}
        [every edge/.style={draw= black, very thick},
        every node/.style={fill= white,circle}]
        \path (A) edge[red] (B);
        \path (A) edge[gray] (C);
        \path (B) edge[red] (C);
        \path (B) edge[red] (D);
        \path (B) edge[red] (E);
        \path (C) edge[gray] (D);
        \path (C) edge[gray] (E);
        \path (D) edge (E);
        \path (D) edge (F);
        \path (E) edge (F);
      \end{scope}
    \end{tikzpicture}
  \end{center}

  (Step 5): By \cref{rmk:graph-exploration-algorithms-main-idea-4}, pick the next vertex $D$.
  We need to move $F$ to the frontier pile and remove $D$ from the frontier pile.
  \begin{center}
    \begin{tikzpicture}[node distance= 2cm]
      \begin{scope}[every node/.style={circle,thick,draw}]
        \node[gray] (A) {$A$};
        \node[gray] (B) [above right of= A] {$B$};
        \node[gray] (C) [below right of= A] {$C$};
        \node[gray] (D) [right of= B] {$D$};
        \node[blue] (E) [right of= C] {$E$};
        \node[blue] (F) [below right of= D] {$F$};
      \end{scope}
      \begin{scope}
        [every edge/.style={draw= black, very thick},
        every node/.style={fill= white,circle}]
        \path (A) edge[red] (B);
        \path (A) edge[gray] (C);
        \path (B) edge[red] (C);
        \path (B) edge[red] (D);
        \path (B) edge[red] (E);
        \path (C) edge[gray] (D);
        \path (C) edge[gray] (E);
        \path (D) edge[gray] (E);
        \path (D) edge[red] (F);
        \path (E) edge (F);
      \end{scope}
    \end{tikzpicture}
  \end{center}

  (Step 6): The unexplored pile is empty.
  By \cref{rmk:graph-exploration-algorithms-main-idea-2}, we can remove everything from the frontier pile.
  \begin{center}
    \begin{tikzpicture}[node distance= 2cm]
      \begin{scope}[every node/.style={circle,thick,draw}]
        \node[gray] (A) {$A$};
        \node[gray] (B) [above right of= A] {$B$};
        \node[gray] (C) [below right of= A] {$C$};
        \node[gray] (D) [right of= B] {$D$};
        \node[gray] (E) [right of= C] {$E$};
        \node[gray] (F) [below right of= D] {$F$};
      \end{scope}
      \begin{scope}
        [every edge/.style={draw= black, very thick},
        every node/.style={fill= white,circle}]
        \path (A) edge[red] (B);
        \path (A) edge[gray] (C);
        \path (B) edge[red] (C);
        \path (B) edge[red] (D);
        \path (B) edge[red] (E);
        \path (C) edge[gray] (D);
        \path (C) edge[gray] (E);
        \path (D) edge[gray] (E);
        \path (D) edge[red] (F);
        \path (E) edge[gray] (F);
      \end{scope}
    \end{tikzpicture}
  \end{center}
\end{eg}

>>>>

\para
\label{para:edges-are-binary-relations}
Recall that a binary relation on a set $V$ can be encoded as any subset of $V \times V$.
Thus, a graph is a set $V$ equipped with a binary relation $E$.

@@

\para
\cref{eg:graphs-example1} is not an undirected graph since $(B,C) \in E$, but $(C,B) \notin E$.
Similarly, $(C,A) \in E$, but $(A,C) \notin E$.

\para
In an undirected graph, we drop the arrow tips as they convey no additional information.

@@

\begin{thm}
  \label{thm:reachability-is-an-equivalence}
  Let $G = (V, E)$ be an undirected graph.
  The s-t reachability relation defined in \cref{defn:reachability} is an equivalence relation on the set of vertices $V$.
\end{thm}
\begin{proof}\hfill
  \para
  Reflexivity: Let $v \in V$.
  The empty walk is trivially a path.
  Thus, $v$ is s-t reachable from itself.

  \para
  Symmetry: Let $u,v \in V$ be any two vertices.
  Assume that $v$ is s-t reachable from $u$, i.e., there is a path $p$ from $u$ to $v$.
  We need to prove that $u$ is s-t reachable from $v$, i.e., there is a path from $v$ to $u$.
  To this end, we do induction on the length of $p$.
  In the base case, $p$ has length 0, so it is the empty path.
  Thus, the result follows from reflexivity.
  In the induction step, $p$ is a path of length $k+1$ and it is of the form $p',(u',v),v$, where $u'$ is some vertex and $p'$ is a path from $u$ to $u'$ of length $k$.
  By the induction hypothesis, there is a path $\widetilde{p}$ from $u'$ to $u$.
  Since $G$ is undirected, $(v,u')$ is also an edge of $G$.
  Thus, the sequence
  \[
    v,(v,u'),\widetilde{p}
  \]
  defines a path from $v$ to $u$.

  \para
  Transitivity: Let $u,v,w \in V$ be any three vertices.
  Assume that there is a path $p$ from $u$ to $v$ and a path $p'$ from $v$ to $w$.
  The concatenation of $p$ and $p'$ defines a path from $u$ to $w$.
  You can prove this rigorously by induction on the length of $p$.
  The detail is left as an exercise.
\end{proof}

@@

\para
\label{para:graph-exploration-algorithms-main-idea-1}
To explore every vertex of an undirected graph $(V,E)$, we divide $V$ into three piles: explored, frontier, and unexplored.
Initially, the unexplored pile contains every vertex and the explored and frontier piles are empty.

\para
\label{para:graph-exploration-algorithms-main-idea-5}
We can explore the graph recursively.
We match the current piles against the conditions stated in
\ref{para:graph-exploration-algorithms-main-idea-2},
\ref{para:graph-exploration-algorithms-main-idea-3}, and
\ref{para:graph-exploration-algorithms-main-idea-4}
in this order.

\para
\label{para:graph-exploration-algorithms-main-idea-2}
If the unexplored pile is empty, we can move every vertex in the frontier pile to the explored pile.
This is the base case of the recursion.

\para
\label{para:graph-exploration-algorithms-main-idea-3}
If the frontier pile is empty, move an arbitrary vertex in the unexplored pile to the frontier pile.
Make a recursive call on the current piles (i.e., go back to \ref{para:graph-exploration-algorithms-main-idea-5}).

\para
\label{para:graph-exploration-algorithms-main-idea-4}
If the frontier pile is not empty, pick the next vertex $v$ in the frontier pile.
If $v$ is not adjacent to any unexplored vertex, then $v$ has been fully explored, so we can move it to the explore pile.
Otherwise, we move an arbitrary vertex adjacent to $v$ in the unexplored pile to the frontier pile.
Finally, we make a recursive call on the current piles (i.e., go back to \ref{para:graph-exploration-algorithms-main-idea-5}).

\begin{rmk}
  \label{rmk:graph-exploration-algorithms-next-vertex}
  The ``next vertex'' in \ref{para:graph-exploration-algorithms-main-idea-4} is kept abstract intentionally.
  There are two natural ways to pick the ``next vertex:'' pick the first vertex or the last vertex added to the frontier pile.
  The former choice corresponds to breadth-first search (BFS) and the latter choice corresponds to depth-first search (DFS).
\end{rmk}

\begin{rmk}
  \label{rmk:graph-exploration-algorithms-termination}
  Each recursive call in
  \ref{para:graph-exploration-algorithms-main-idea-3} and
  \ref{para:graph-exploration-algorithms-main-idea-4}
  is made on a smaller unexplored pile.
  We can prove theorems about this algorithm by strong induction on the size of the unexplored pile.
\end{rmk}

\begin{eg}
  \label{eg:bfs-example}
  Run BFS on the following graph:
  \begin{center}
    \begin{tikzpicture}[node distance= 2cm]
      \begin{scope}[every node/.style={circle,thick,draw}]
        \node (A) {$A$};
        \node (B) [above right of= A] {$B$};
        \node (C) [below right of= A] {$C$};
        \node (D) [right of= B] {$D$};
        \node (E) [right of= C] {$E$};
        \node (F) [below right of= D] {$F$};
      \end{scope}
      \begin{scope}
        [every edge/.style={draw= black, very thick},
        every node/.style={fill= white,circle}]
        \path (A) edge (B);
        \path (A) edge (C);
        \path (B) edge (D);
        \path (C) edge (D);
        \path (D) edge (E);
        \path (D) edge (F);
        \path (E) edge (F);
      \end{scope}
    \end{tikzpicture}
  \end{center}

  \para
  Initially, the frontier pile is empty.
  This matches \ref{para:graph-exploration-algorithms-main-idea-3}, so we move an arbitrary vertex in the unexplored pile to the frontier pile.
  \begin{center}
    \begin{tikzpicture}[node distance= 2cm]
      \begin{scope}[every node/.style={circle,thick,draw}]
        \node (A) {$A$};
        \node (B)[blue] [above right of= A] {$B$};
        \node (C) [below right of= A] {$C$};
        \node (D) [right of= B] {$D$};
        \node (E) [right of= C] {$E$};
        \node (F) [below right of= D] {$F$};
      \end{scope}
      \begin{scope}
        [every edge/.style={draw= black, very thick},
        every node/.style={fill= white,circle}]
        \path (A) edge (B);
        \path (A) edge (C);
        \path (B) edge (D);
        \path (C) edge (D);
        \path (D) edge (E);
        \path (D) edge (F);
        \path (E) edge (F);
      \end{scope}
    \end{tikzpicture}
  \end{center}

  \para
  The frontier pile is not empty.
  This matches \ref{para:graph-exploration-algorithms-main-idea-4}.
  $B$ is adjacent to an explored vertex (e.g., $A$) so we move $A$ to the frontier pile.
  \begin{center}
    \begin{tikzpicture}[node distance= 2cm]
      \begin{scope}[every node/.style={circle,thick,draw}]
        \node (A)[blue] {$A$};
        \node (B)[blue] [above right of= A] {$B$};
        \node (C) [below right of= A] {$C$};
        \node (D) [right of= B] {$D$};
        \node (E) [right of= C] {$E$};
        \node (F) [below right of= D] {$F$};
      \end{scope}
      \begin{scope}
        [every edge/.style={draw= black, very thick},
        every node/.style={fill= white,circle}]
        \path (A) edge[red] (B);
        \path (A) edge (C);
        \path (B) edge (D);
        \path (C) edge (D);
        \path (D) edge (E);
        \path (D) edge (F);
        \path (E) edge (F);
      \end{scope}
    \end{tikzpicture}
  \end{center}

  \para
  The frontier pile is not empty.
  This matches \ref{para:graph-exploration-algorithms-main-idea-4}.
  $B$ is added to the frontier pile first, so we choose to explore it.
  \begin{center}
    \begin{tikzpicture}[node distance= 2cm]
      \begin{scope}[every node/.style={circle,thick,draw}]
        \node (A)[blue] {$A$};
        \node (B)[blue] [above right of= A] {$B$};
        \node (C) [below right of= A] {$C$};
        \node (D)[blue] [right of= B] {$D$};
        \node (E) [right of= C] {$E$};
        \node (F) [below right of= D] {$F$};
      \end{scope}
      \begin{scope}
        [every edge/.style={draw= black, very thick},
        every node/.style={fill= white,circle}]
        \path (A) edge[red] (B);
        \path (A) edge (C);
        \path (B) edge[red] (D);
        \path (C) edge (D);
        \path (D) edge (E);
        \path (D) edge (F);
        \path (E) edge (F);
      \end{scope}
    \end{tikzpicture}
  \end{center}

  \para
  The frontier pile is not empty.
  This matches \ref{para:graph-exploration-algorithms-main-idea-4}.
  Since $B$ is not adjacent to any unexplored vertex, we can move it to the explored pile.
  \begin{center}
    \begin{tikzpicture}[node distance= 2cm]
      \begin{scope}[every node/.style={circle,thick,draw}]
        \node (A)[blue] {$A$};
        \node (B)[gray] [above right of= A] {$B$};
        \node (C) [below right of= A] {$C$};
        \node (D)[blue] [right of= B] {$D$};
        \node (E) [right of= C] {$E$};
        \node (F) [below right of= D] {$F$};
      \end{scope}
      \begin{scope}
        [every edge/.style={draw= black, very thick},
        every node/.style={fill= white,circle}]
        \path (A) edge[red] (B);
        \path (A) edge (C);
        \path (B) edge[red] (D);
        \path (C) edge (D);
        \path (D) edge (E);
        \path (D) edge (F);
        \path (E) edge (F);
      \end{scope}
    \end{tikzpicture}
  \end{center}

  %

  \begin{center}
    \begin{tikzpicture}[node distance= 2cm]
      \begin{scope}[every node/.style={circle,thick,draw}]
        \node (A)[blue] {$A$};
        \node (B)[gray] [above right of= A] {$B$};
        \node (C)[blue] [below right of= A] {$C$};
        \node (D)[blue] [right of= B] {$D$};
        \node (E) [right of= C] {$E$};
        \node (F) [below right of= D] {$F$};
      \end{scope}
      \begin{scope}
        [every edge/.style={draw= black, very thick},
        every node/.style={fill= white,circle}]
        \path (A) edge[red] (B);
        \path (A) edge[red] (C);
        \path (B) edge[red] (D);
        \path (C) edge (D);
        \path (D) edge (E);
        \path (D) edge (F);
        \path (E) edge (F);
      \end{scope}
    \end{tikzpicture}
  \end{center}

  %

  \begin{center}
    \begin{tikzpicture}[node distance= 2cm]
      \begin{scope}[every node/.style={circle,thick,draw}]
        \node (A)[gray] {$A$};
        \node (B)[gray] [above right of= A] {$B$};
        \node (C)[blue] [below right of= A] {$C$};
        \node (D)[blue] [right of= B] {$D$};
        \node (E) [right of= C] {$E$};
        \node (F) [below right of= D] {$F$};
      \end{scope}
      \begin{scope}
        [every edge/.style={draw= black, very thick},
        every node/.style={fill= white,circle}]
        \path (A) edge[red] (B);
        \path (A) edge[red] (C);
        \path (B) edge[red] (D);
        \path (C) edge (D);
        \path (D) edge (E);
        \path (D) edge (F);
        \path (E) edge (F);
      \end{scope}
    \end{tikzpicture}
  \end{center}

  %

  \begin{center}
    \begin{tikzpicture}[node distance= 2cm]
      \begin{scope}[every node/.style={circle,thick,draw}]
        \node (A)[gray] {$A$};
        \node (B)[gray] [above right of= A] {$B$};
        \node (C)[blue] [below right of= A] {$C$};
        \node (D)[blue] [right of= B] {$D$};
        \node (E)[blue] [right of= C] {$E$};
        \node (F) [below right of= D] {$F$};
      \end{scope}
      \begin{scope}
        [every edge/.style={draw= black, very thick},
        every node/.style={fill= white,circle}]
        \path (A) edge[red] (B);
        \path (A) edge[red] (C);
        \path (B) edge[red] (D);
        \path (C) edge (D);
        \path (D) edge[red] (E);
        \path (D) edge (F);
        \path (E) edge (F);
      \end{scope}
    \end{tikzpicture}
  \end{center}

  %

  \begin{center}
    \begin{tikzpicture}[node distance= 2cm]
      \begin{scope}[every node/.style={circle,thick,draw}]
        \node (A)[gray] {$A$};
        \node (B)[gray] [above right of= A] {$B$};
        \node (C)[blue] [below right of= A] {$C$};
        \node (D)[blue] [right of= B] {$D$};
        \node (E)[blue] [right of= C] {$E$};
        \node (F)[blue] [below right of= D] {$F$};
      \end{scope}
      \begin{scope}
        [every edge/.style={draw= black, very thick},
        every node/.style={fill= white,circle}]
        \path (A) edge[red] (B);
        \path (A) edge[red] (C);
        \path (B) edge[red] (D);
        \path (C) edge (D);
        \path (D) edge[red] (E);
        \path (D) edge[red] (F);
        \path (E) edge (F);
      \end{scope}
    \end{tikzpicture}
  \end{center}

  \para
  At this point, the unexplored pile is empty, so we have reached the base case \ref{para:graph-exploration-algorithms-main-idea-2}.
  \begin{center}
    \begin{tikzpicture}[node distance= 2cm]
      \begin{scope}[every node/.style={circle,thick,draw}]
        \node (A)[gray] {$A$};
        \node (B)[gray] [above right of= A] {$B$};
        \node (C)[gray] [below right of= A] {$C$};
        \node (D)[gray] [right of= B] {$D$};
        \node (E)[gray] [right of= C] {$E$};
        \node (F)[gray] [below right of= D] {$F$};
      \end{scope}
      \begin{scope}
        [every edge/.style={draw= black, very thick},
        every node/.style={fill= white,circle}]
        \path (A) edge[red] (B);
        \path (A) edge[red] (C);
        \path (B) edge[red] (D);
        \path (C) edge[gray] (D);
        \path (D) edge[red] (E);
        \path (D) edge[red] (F);
        \path (E) edge[gray] (F);
      \end{scope}
    \end{tikzpicture}
  \end{center}
\end{eg}

\begin{eg}
  \label{eg:dfs-example}
  Run DFS on the same graph defined in \cref{eg:bfs-example}.
  \begin{center}
    \begin{tikzpicture}[node distance= 2cm]
      \begin{scope}[every node/.style={circle,thick,draw}]
        \node (A) {$A$};
        \node (B) [above right of= A] {$B$};
        \node (C) [below right of= A] {$C$};
        \node (D) [right of= B] {$D$};
        \node (E) [right of= C] {$E$};
        \node (F) [below right of= D] {$F$};
      \end{scope}
      \begin{scope}
        [every edge/.style={draw= black, very thick},
        every node/.style={fill= white,circle}]
        \path (A) edge (B);
        \path (A) edge (C);
        \path (B) edge (D);
        \path (C) edge (D);
        \path (D) edge (E);
        \path (D) edge (F);
        \path (E) edge (F);
      \end{scope}
    \end{tikzpicture}
  \end{center}

  \para
  This time, let us arbitrarily pick the vertex $C$.
  \begin{center}
    \begin{tikzpicture}[node distance= 2cm]
      \begin{scope}[every node/.style={circle,thick,draw}]
        \node (A) {$A$};
        \node (B) [above right of= A] {$B$};
        \node (C)[blue] [below right of= A] {$C$};
        \node (D) [right of= B] {$D$};
        \node (E) [right of= C] {$E$};
        \node (F) [below right of= D] {$F$};
      \end{scope}
      \begin{scope}
        [every edge/.style={draw= black, very thick},
        every node/.style={fill= white,circle}]
        \path (A) edge (B);
        \path (A) edge (C);
        \path (B) edge (D);
        \path (C) edge (D);
        \path (D) edge (E);
        \path (D) edge (F);
        \path (E) edge (F);
      \end{scope}
    \end{tikzpicture}
  \end{center}

  %

  \begin{center}
    \begin{tikzpicture}[node distance= 2cm]
      \begin{scope}[every node/.style={circle,thick,draw}]
        \node (A) {$A$};
        \node (B) [above right of= A] {$B$};
        \node (C)[blue] [below right of= A] {$C$};
        \node (D)[blue] [right of= B] {$D$};
        \node (E) [right of= C] {$E$};
        \node (F) [below right of= D] {$F$};
      \end{scope}
      \begin{scope}
        [every edge/.style={draw= black, very thick},
        every node/.style={fill= white,circle}]
        \path (A) edge (B);
        \path (A) edge (C);
        \path (B) edge (D);
        \path (C) edge[red] (D);
        \path (D) edge (E);
        \path (D) edge (F);
        \path (E) edge (F);
      \end{scope}
    \end{tikzpicture}
  \end{center}

  \para
  Since $D$ is added after $C$, we need to explore $D$ first.
  \begin{center}
    \begin{tikzpicture}[node distance= 2cm]
      \begin{scope}[every node/.style={circle,thick,draw}]
        \node (A) {$A$};
        \node (B) [above right of= A] {$B$};
        \node (C)[blue] [below right of= A] {$C$};
        \node (D)[blue] [right of= B] {$D$};
        \node (E)[blue] [right of= C] {$E$};
        \node (F) [below right of= D] {$F$};
      \end{scope}
      \begin{scope}
        [every edge/.style={draw= black, very thick},
        every node/.style={fill= white,circle}]
        \path (A) edge (B);
        \path (A) edge (C);
        \path (B) edge (D);
        \path (C) edge[red] (D);
        \path (D) edge[red] (E);
        \path (D) edge (F);
        \path (E) edge (F);
      \end{scope}
    \end{tikzpicture}
  \end{center}

  %

  \begin{center}
    \begin{tikzpicture}[node distance= 2cm]
      \begin{scope}[every node/.style={circle,thick,draw}]
        \node (A) {$A$};
        \node (B) [above right of= A] {$B$};
        \node (C)[blue] [below right of= A] {$C$};
        \node (D)[blue] [right of= B] {$D$};
        \node (E)[blue] [right of= C] {$E$};
        \node (F)[blue] [below right of= D] {$F$};
      \end{scope}
      \begin{scope}
        [every edge/.style={draw= black, very thick},
        every node/.style={fill= white,circle}]
        \path (A) edge (B);
        \path (A) edge (C);
        \path (B) edge (D);
        \path (C) edge[red] (D);
        \path (D) edge[red] (E);
        \path (D) edge (F);
        \path (E) edge[red] (F);
      \end{scope}
    \end{tikzpicture}
  \end{center}

  % 
  \para
  At this point, $F$ has been fully explored.
  We backtrack until we find a vertex that has not been fully explored.
  \begin{center}
    \begin{tikzpicture}[node distance= 2cm]
      \begin{scope}[every node/.style={circle,thick,draw}]
        \node (A) {$A$};
        \node (B) [above right of= A] {$B$};
        \node (C)[blue] [below right of= A] {$C$};
        \node (D)[blue] [right of= B] {$D$};
        \node (E)[blue] [right of= C] {$E$};
        \node (F)[gray] [below right of= D] {$F$};
      \end{scope}
      \begin{scope}
        [every edge/.style={draw= black, very thick},
        every node/.style={fill= white,circle}]
        \path (A) edge (B);
        \path (A) edge (C);
        \path (B) edge (D);
        \path (C) edge[red] (D);
        \path (D) edge[red] (E);
        \path (D) edge[gray] (F);
        \path (E) edge[red] (F);
      \end{scope}
    \end{tikzpicture}
  \end{center}

  %

  \begin{center}
    \begin{tikzpicture}[node distance= 2cm]
      \begin{scope}[every node/.style={circle,thick,draw}]
        \node (A) {$A$};
        \node (B) [above right of= A] {$B$};
        \node (C)[blue] [below right of= A] {$C$};
        \node (D)[blue] [right of= B] {$D$};
        \node (E)[gray] [right of= C] {$E$};
        \node (F)[gray] [below right of= D] {$F$};
      \end{scope}
      \begin{scope}
        [every edge/.style={draw= black, very thick},
        every node/.style={fill= white,circle}]
        \path (A) edge (B);
        \path (A) edge (C);
        \path (B) edge (D);
        \path (C) edge[red] (D);
        \path (D) edge[red] (E);
        \path (D) edge[gray] (F);
        \path (E) edge[red] (F);
      \end{scope}
    \end{tikzpicture}
  \end{center}

  %

  \begin{center}
    \begin{tikzpicture}[node distance= 2cm]
      \begin{scope}[every node/.style={circle,thick,draw}]
        \node (A) {$A$};
        \node (B)[blue] [above right of= A] {$B$};
        \node (C)[blue] [below right of= A] {$C$};
        \node (D)[blue] [right of= B] {$D$};
        \node (E)[gray] [right of= C] {$E$};
        \node (F)[gray] [below right of= D] {$F$};
      \end{scope}
      \begin{scope}
        [every edge/.style={draw= black, very thick},
        every node/.style={fill= white,circle}]
        \path (A) edge (B);
        \path (A) edge (C);
        \path (B) edge[red] (D);
        \path (C) edge[red] (D);
        \path (D) edge[red] (E);
        \path (D) edge[gray] (F);
        \path (E) edge[red] (F);
      \end{scope}
    \end{tikzpicture}
  \end{center}

  %

  \begin{center}
    \begin{tikzpicture}[node distance= 2cm]
      \begin{scope}[every node/.style={circle,thick,draw}]
        \node (A)[blue] {$A$};
        \node (B)[blue] [above right of= A] {$B$};
        \node (C)[blue] [below right of= A] {$C$};
        \node (D)[blue] [right of= B] {$D$};
        \node (E)[gray] [right of= C] {$E$};
        \node (F)[gray] [below right of= D] {$F$};
      \end{scope}
      \begin{scope}
        [every edge/.style={draw= black, very thick},
        every node/.style={fill= white,circle}]
        \path (A) edge[red] (B);
        \path (A) edge (C);
        \path (B) edge[red] (D);
        \path (C) edge[red] (D);
        \path (D) edge[red] (E);
        \path (D) edge[gray] (F);
        \path (E) edge[red] (F);
      \end{scope}
    \end{tikzpicture}
  \end{center}

  %

  \begin{center}
    \begin{tikzpicture}[node distance= 2cm]
      \begin{scope}[every node/.style={circle,thick,draw}]
        \node (A)[gray] {$A$};
        \node (B)[gray] [above right of= A] {$B$};
        \node (C)[gray] [below right of= A] {$C$};
        \node (D)[gray] [right of= B] {$D$};
        \node (E)[gray] [right of= C] {$E$};
        \node (F)[gray] [below right of= D] {$F$};
      \end{scope}
      \begin{scope}
        [every edge/.style={draw= black, very thick},
        every node/.style={fill= white,circle}]
        \path (A) edge[red] (B);
        \path (A) edge[gray] (C);
        \path (B) edge[red] (D);
        \path (C) edge[red] (D);
        \path (D) edge[red] (E);
        \path (D) edge[gray] (F);
        \path (E) edge[red] (F);
      \end{scope}
    \end{tikzpicture}
  \end{center}
\end{eg}
